ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-09-30 rebuttal.tex

\begin{response}
    %
    The reviewer is correct, and this was a typo in the abstract. We have clarified that PyKED is
    intended as software to validate and work with ChemKED files.

    That said, we do have ongoing work to create a graphical user interface to create ChemKED files,
    as this was requested by several of the experimental teams that have expressed interest in the
    ChemKED format. The work is available publicly at
    \url{https://github.com/pr-omethe-us/ChemKED-gui}.
    %
\end{response}

\begin{reviewer}
    %
    5.~Page 1, line 31: The future development plans include support for jet stirred reactors,
    however, this is not mentioned in section 5, Conclusions and future work, nor is there any
    reference to it on the GitHub wiki roadmap of PyKED.
    %
\end{reviewer}

\begin{response}
    %
    Thank you for the comment, we have corrected this oversight in the roadmap and the conclusions.
    %
\end{response}

@@

\begin{reviewer}
    %
    6. I don't completely agree with the second part of the statement on page 2 line 16-18:
    ``Second, the XML format is intended to be a machine-readable mark-up language rather than a
    data format, and its lack of human readability presents a barrier to creating and working with
    database files.''

    In my opinion the YAML format resembles the XML format strongly. XML format is human-readable,
    however, because of the html-tags and \texttt{<>} characters, the YAML format might be easier to
    read. If XML would lack human readability as it is stated in the article, it would not be such a
    popular format, and I don't think PrIMe and ReSpecTh would be using it.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that XML is technically human readable by an expert (as opposed to,
    say, a binary format), and also that YAML is easier to read---by ``human-readable'', we really
    mean human-\emph{understandable}. We have modified the text to clarify this point; please see
    also the response to a similar comment from Reviewer \#1.
    %
\end{response}

### 2017-09-30 rebuttal.tex

\begin{response}
    %
    We have edited that sentence to correct the error.
    %
\end{response}

\subsection*{Overview of ChemKED format}

\begin{reviewer}
    %
    9.~A reference should not be a required mapping of an experimental dataset. If the intention is
    to make the community to adopt the ChemKED file structure, it might be easier to do so if
    experimentalists can use the structure internally as well, i.e. to read, store, and share
    non-published internal data.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that some of the choices currently enforced in the PyKED package and
    the ChemKED schema do not lend themselves to easy informal use. We presently recommend that
    users who wish to use ChemKED files in this way include an incomplete reference field since the
    only required subfields of the reference field are the authors and the year.

    In addition, there is a \verb|skip_validation| argument to the \verb|ChemKED| class constructor.
    This option turns off all of the validation, including the validation of the reference material
    and enforcement of required fields. We recognize that this is a limitation at present, if users
    wish to, for example, validate that their units are correct but do not wish to validate some
    other aspect of the file. Work is ongoing to improve this situation, and discussion is contained
    in a public issue on GitHub: \url{https://github.com/pr-omethe-us/PyKED/issues/39}
    %
\end{response}

### 2017-09-30 rebuttal.tex

\begin{response}

\end{response}

\begin{reviewer}
    %
    13.~Page 6, line 12 describes the initial efforts of building an open repository of ChemKED
    files, on GitHub. Is this the most efficient way for making a publically available database? One
    issue with this is the fact that adding files is done via pull requests, which have to be
    verified and approved by the owners of the repository. Therefore a certain delay is inevitable.
    %
\end{reviewer}

\begin{response}
    %
    We agree that submitting files to the ChemKED database on GitHub will inevitably incur some
    delay, however we see the human intervention step as necessary and important. In much the same
    way that review of manuscripts provides a certain surety of the minimal quality of the article,
    we see review of data being added to the database as a minimal quality check; for instance, does
    the reference exist, has the data already been added to the database, are the files styled
    consistently, etc. Furthermore, such human review can help to catch some errors that may be
    introduced by manual creation of the files, as pointed out by the reviewer previously. Finally,
    we would note that all of the existing database have a similar requirement, where users submit
    data files for inclusion that are approved by the database owners after some review. In this
    sense, we actually see the Pull Request mechanism of GitHub as an advantage, since all reviews
    can be done publicly and some of the review procedures (style checks, reference validity, etc)
    may be done automatically by a script associated with the GitHub repository. In this way, the
    entire community can contribute to improving the quality of the database.
    %
\end{response}

### 2017-09-30 rebuttal.tex

\begin{reviewer}
    %
    13.~Page 6, line 12 describes the initial efforts of building an open repository of ChemKED
    files, on GitHub. Is this the most efficient way for making a publically available database? One
    issue with this is the fact that adding files is done via pull requests, which have to be
    verified and approved by the owners of the repository. Therefore a certain delay is inevitable.
    %
\end{reviewer}

\begin{response}
    %
    We agree that submitting files to the ChemKED database on GitHub will inevitably incur some
    delay, however we see the human intervention step as necessary and important. In much the same
    way that review of manuscripts provides a certain surety of the minimal quality of the article,
    we see review of data being added to the database as a minimal quality check; for instance, does
    the reference exist, has the data already been added to the database, are the files styled
    consistently, etc. Furthermore, such human review can help to catch some errors that may be
    introduced by manual creation of the files, as pointed out by the reviewer previously. Finally,
    we would note that all of the existing database have a similar requirement, where users submit
    data files for inclusion that are approved by the database owners after some review. In this
    sense, we actually see the Pull Request mechanism of GitHub as an advantage, since all reviews
    can be done publicly and some of the review procedures (style checks, reference validity, etc)
    may be done automatically by a script associated with the GitHub repository. In this way, the
    entire community can contribute to improving the quality of the database.
    %
\end{response}

\begin{reviewer}
    %
    14.~Page 6, line 18 states that a converter exists from ReSpecTh to ChemKED. But isn't this
    feature added when going from v0.1.6 to v0.2.0, which makes page 3 line 21: v0.1.6 not correct.
    %
\end{reviewer}

\begin{response}
    %
    In principle, the version of the ChemKED database format and the version of the PyKED Python
    package are entirely independent. Thus, since adding the converters incurred no changes to the
    database format, the description of the format in the manuscript is a description of v0.1.6,
    which happens to be identical to v0.2.0. However, in practice, the versions of PyKED and the
    ChemKED database have been coupled up to now, and we have updated the version mentioned in the
    text to vX.X.X, which we released in conjunction with submitting this response after the helpful
    comments of the reviewers.
    %
\end{response}

### 2017-09-30 rebuttal.tex

\begin{response}
    %
    We removed the extraneous comma.
    %
\end{response}

\subsection*{PyKED architecture}

\begin{reviewer}
    %
    16.~The Python package Cerberus is used to validate the format and content of the given ChemKED
    file. How does this package handles/reports the inconsistencies with the schemas? As far as my
    local test went with giving a corrupted YAML file to the ChemKED initiator, the package just
    spits out a Python error, which might be not the most user-friendly solution.
    %
\end{reviewer}

\begin{response}
    %
    We recognize that this is a limitation of the current version of PyKED. The technical reason for
    this in the Cerberus package has very recently been resolved (see
    \url{https://github.com/pyeve/cerberus/issues/278}), so we plan to update the error reporting
    mechanism in PyKED to report well-formatted errors that have a detailed description of the
    error. We have opened a public issue to track the progress on this:
    \url{https://github.com/pr-omethe-us/PyKED/issues/78}. Thank you for suggesting this
    improvement.
    %
\end{response}

### 2017-10-01 rebuttal.tex

\begin{response}
    %
    We have edited that sentence to correct the error.
    %
\end{response}

\subsection*{Overview of ChemKED format}

\begin{reviewer}
    %
    9.~A reference should not be a required mapping of an experimental dataset. If the intention is
    to make the community to adopt the ChemKED file structure, it might be easier to do so if
    experimentalists can use the structure internally as well, i.e. to read, store, and share
    non-published internal data.
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that some of the choices currently enforced in the PyKED package and
    the ChemKED schema do not lend themselves to easy informal use. We presently recommend that
    users who wish to use ChemKED files in this way include an incomplete reference field. We have
    modified the ChemKED schema in response to this comment so that the only required subfields of
    the reference field are the authors and the year.

### 2017-10-01 rebuttal.tex

\begin{reviewer}
    %
    16.~The Python package Cerberus is used to validate the format and content of the given ChemKED
    file. How does this package handles/reports the inconsistencies with the schemas? As far as my
    local test went with giving a corrupted YAML file to the ChemKED initiator, the package just
    spits out a Python error, which might be not the most user-friendly solution.
    %
\end{reviewer}

\begin{response}
    %
    We recognize that this is a limitation of the current version of PyKED. The technical reason for
    this in the Cerberus package has very recently been resolved (see
    \url{https://github.com/pyeve/cerberus/issues/278}), so we plan to update the error reporting
    mechanism in PyKED to report well-formatted errors that have a detailed description of the
    error. We have opened a public issue to track the progress on this:
    \url{https://github.com/pr-omethe-us/PyKED/issues/78}. Thank you for suggesting this
    improvement.
    %
\end{response}

\begin{reviewer}
    %
    17.~A recurring issue with actively developing codes is backwards compatibility. How are the
    authors planning on addressing this issue?
    %
\end{reviewer}

\begin{response}
    %
    We agree with the reviewer that backwards compatibility may become an issue. We have addressed
    this concern in ChemKED by including the database version as one of the required keys in the
    ChemKED file. Although the functionality doesn't exist yet, we intend to be able to check this
    version and report any feature deprecation or removal to the user (see
    \url{https://github.com/pr-omethe-us/PyKED/issues/22), and we plan to provide an automated
    upgrade path for ChemKED files when backwards-incompatible changes are made. In addition, we use
    the widely supported Semantic Versioning paradigm, where changes in the various parts of the
    version number (vX.Y.Z) have specific meaning; changing the major version (X) indicates a
    backwards incompatible change, changing the minor version (Y) indicates a backwards compatible,
    but somewhat large, change, and changing the patch version (Z) indicates a bug fix with no
    anticipated incompatibilities.

    Finally, each released version of PyKED is permanently archived on Zenodo
    (\url{https://doi.org/10.5281/zenodo.597935}) so that it is accessible should a user wish to use
    an old version to solve a compatibility issue temporarily.
    %
\end{response}

### 2017-10-02 pyked.tex
Fix response to #26
Turns out the journal was also required, so make a note that we've
dropped that requirement. Also, fix the text where some optional fields
were specified as required.
Respond to #35
Correct the typo in the manuscript regarding hyphens in composition type
names

The \yabox|value| can be a string, integer, or floating-point number. We designed
the \ck{} format to include all information necessary to simulate a
given experiment. A \ck{} file is generally broken into two main sections.
The first section details all the ``meta'' information about the experiment
and the \ck{} file itself. This section uses the common \yabox|author| mapping
for any fields that specify an author. The \yabox|author| mapping
contains the following fields:
%
\begin{itemize}
    \item \yabox|name| (required, string): The author's full name
    \item \yabox|ORCID| (optional, string): the author's unique ORCID code
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
%
\begin{itemize}
    \item \yabox|file-author| (required, \yabox|author|-type mapping): The
    author of the \ck{} file
    \item \yabox|file-version| (required, integer): The version of the \ck{}
    file
    \item \yabox|chemked-version| (required, string): The version of the \ck{}
    schema to which this file conforms
    \item \yabox|reference| (required, mapping): The reference information for
    the published article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yabox|authors| (required, sequence): A sequence of \yabox|author|
        mappings
        \item \yabox|journal| (required, string): The name of the publishing
        journal
        \item \yabox|year| (required, integer): The year of publication
        \item \yabox|doi| (optional, string): The article DOI
        \item \yabox|volume| (required, integer): The journal volume number
        \item \yabox|pages| (optional, string): The article pages
        \item \yabox|detail| (optional, string): A description of from where the
        data originated (e.g., figure or table number)
    \end{itemize}
    \item \yabox|experiment-type| (required, string): Currently, only
    \yabox|ignition delay| is supported
    \item \yabox|apparatus| (required, string): The type of apparatus used to
    perform the experiment (\yabox|shock tube| or \yabox|rapid compression machine|)
\end{itemize}

@@

The \pybox|DataPoint| class stores the information associated with a single data
point in the \ck{} file (i.e., a single element of the \yabox|datapoints|
sequence). Similar to the \pybox|ChemKED| class,
the \pybox|DataPoint| stores information as instance attributes:
%
\begin{itemize}
    \item \pybox|equivalence_ratio|: The value of the equivalence ratio, if
    present. For informational purposes only---no validation is done of the
    value.
    \item \pybox|ignition_delay|, \pybox|temperature|, \pybox|pressure|,
    \pybox|pressure_rise|: The value of these data are stored as instances of
    the \pybox|Quantity| class from the Pint \autocite{Grecco2016} package,
    possibly with associated uncertainty
    \item \pybox|composition|: A list of dictionaries of the species and their
    respective amounts. The values are validated so that \yabox|mole-percent|,
    \yabox|mole-fraction|, or \yabox|mass-fraction| cannot be mixed for a single
    data point, and so that the sum of the values is approximately 1.0, or
    100.0 for \yabox|mole-percent|.
    \item \pybox|composition_type|: A string indicating the type of composition
    information for the data point---one of \pybox|'mole-percent'|,
    \pybox|'mole-fraction'|, or \pybox|'mass-fraction'|
    \item \pybox|volume_history|: If the \yabox|volume-history| of an RCM
    experiment is provided in the \ck{} file, it is stored in this attribute as
    a \pybox|namedtuple|, and the actual values are stored in NumPy arrays
    \autocite{vanderWalt:2011np}
    \item \pybox|compression_time|: If the \ck{} file encodes an RCM experiment,
    the length of the compression stroke may be included
    \item \pybox|compressed_temperature|: If the \ck{} file encodes an RCM experiment,
    the temperature at the end of compression may be included
    \item \pybox|compressed_pressure|: If the \ck{} file encodes an RCM experiment,
    the pressure at the end of compression may be included
    \item \pybox|ignition_type|: A dictionary specifying the method of the
    measurement of ignition delay
\end{itemize}

>>>>

The \yabox|value| can be a string, integer, or floating-point number. We designed
the \ck{} format to include all information necessary to simulate a
given experiment. A \ck{} file is generally broken into two main sections.
The first section details all the ``meta'' information about the experiment
and the \ck{} file itself. This section uses the common \yabox|author| mapping
for any fields that specify an author. The \yabox|author| mapping
contains the following fields:
%
\begin{itemize}
    \item \yabox|name| (required, string): The author's full name
    \item \yabox|ORCID| (optional, string): the author's unique ORCID code
\end{itemize}
%
More generally for the meta information, the required keys in this section
are:
%
\begin{itemize}
    \item \yabox|file-author| (required, \yabox|author|-type mapping): The
    author of the \ck{} file
    \item \yabox|file-version| (required, integer): The version of the \ck{}
    file
    \item \yabox|chemked-version| (required, string): The version of the \ck{}
    schema to which this file conforms
    \item \yabox|reference| (required, mapping): The reference information for
    the published article associated with the data in the file. The fields in this
    mapping are:
    \begin{itemize}
        \item \yabox|authors| (required, sequence): A sequence of \yabox|author|
        mappings
        \item \yabox|journal| (optional, string): The name of the publishing
        journal
        \item \yabox|year| (required, integer): The year of publication
        \item \yabox|doi| (optional, string): The article DOI
        \item \yabox|volume| (optional, integer): The journal volume number
        \item \yabox|pages| (optional, string): The article pages
        \item \yabox|detail| (optional, string): A description of from where the
        data originated (e.g., figure or table number)
    \end{itemize}
    \item \yabox|experiment-type| (required, string): Currently, only
    \yabox|ignition delay| is supported
    \item \yabox|apparatus| (required, string): The type of apparatus used to
    perform the experiment (\yabox|shock tube| or \yabox|rapid compression machine|)
\end{itemize}

@@

The \pybox|DataPoint| class stores the information associated with a single data
point in the \ck{} file (i.e., a single element of the \yabox|datapoints|
sequence). Similar to the \pybox|ChemKED| class,
the \pybox|DataPoint| stores information as instance attributes:
%
\begin{itemize}
    \item \pybox|equivalence_ratio|: The value of the equivalence ratio, if
    present. For informational purposes only---no validation is done of the
    value.
    \item \pybox|ignition_delay|, \pybox|temperature|, \pybox|pressure|,
    \pybox|pressure_rise|: The value of these data are stored as instances of
    the \pybox|Quantity| class from the Pint \autocite{Grecco2016} package,
    possibly with associated uncertainty
    \item \pybox|composition|: A list of dictionaries of the species and their
    respective amounts. The values are validated so that \yabox|mole percent|,
    \yabox|mole fraction|, or \yabox|mass fraction| cannot be mixed for a single
    data point, and so that the sum of the values is approximately 1.0, or
    100.0 for \yabox|mole percent|.
    \item \pybox|composition_type|: A string indicating the type of composition
    information for the data point---one of \pybox|'mole percent'|,
    \pybox|'mole fraction'|, or \pybox|'mass fraction'|
    \item \pybox|volume_history|: If the \yabox|volume-history| of an RCM
    experiment is provided in the \ck{} file, it is stored in this attribute as
    a \pybox|namedtuple|, and the actual values are stored in NumPy arrays
    \autocite{vanderWalt:2011np}
    \item \pybox|compression_time|: If the \ck{} file encodes an RCM experiment,
    the length of the compression stroke may be included
    \item \pybox|compressed_temperature|: If the \ck{} file encodes an RCM experiment,
    the temperature at the end of compression may be included
    \item \pybox|compressed_pressure|: If the \ck{} file encodes an RCM experiment,
    the pressure at the end of compression may be included
    \item \pybox|ignition_type|: A dictionary specifying the method of the
    measurement of ignition delay
\end{itemize}
